Append_smoothed_results

### 2016-11-10 Code/append_smoothed_results.Rmd
finish stuff with bayes models

```{r calculate_life_expectancy}
list.life.tables.smoothed <- vector(mode = "list", length = length(non.zero.stratums.smoothed))
list.index <- 1

@@

  list.life.tables.smoothed[[list.index]] <- life.table(data = subset(dat.smoothed.aggregatedCOD, StateYearRaceSex == stratum.id), 
                                               num.ages.in.group = "nx", 
                                               death.counts = "total_smoothed_deaths", 
                                               population.counts = "Population")
  list.index <- list.index + 1
  print(list.index)
  }
)

```

@@

```{r create_dataset_BlackWhite}
#this dataset contains the black-white data within the same row (wide format) for each year-state-sex strata

BlackSmooth <- dat2 %>% filter(Race2 == "Black" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, Pop_across_age, life.expectancy.birth.smoothed) %>%
                  rename(Pop_black = Pop_across_age,  le.smoothed.black = life.expectancy.birth.smoothed)

WhiteSmooth <- dat2 %>% filter(Race2 == "White" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, Pop_across_age, life.expectancy.birth.smoothed) %>%
                  rename(Pop_white = Pop_across_age,  le.smoothed.white = life.expectancy.birth.smoothed)

@@

BlackWhiteSmooth <- BlackWhiteSmooth %>% mutate(WBgap = le.smoothed.white - le.smoothed.black)
```

@@

```{r save_specified_files}
save(dat.clean, dat2, BlackWhite, paired.ids, list.age.decomp.tables, list.cod.decomp.tables,
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/main_datasets.Rdata")
```

>>>>

```{r calculate_life_expectancy}
#6 mins

list.life.tables.smoothed <- vector(mode = "list", length = length(non.zero.stratums.smoothed))
list.index <- 1

@@

  list.life.tables.smoothed[[list.index]] <- life.table(data = subset(dat.smoothed.aggregatedCOD, StateYearRaceSex == stratum.id), 
                                               num.ages.in.group = "nx", 
                                               death.counts = "total_smoothed_deaths", 
                                               population.counts = "Population")
  list.index <- list.index + 1
  print(list.index)
  }
)
```

@@

```{r create_dataset_BlackWhiteSmooth}
#this dataset contains the black-white data within the same row (wide format) for each year-state-sex strata

BlackSmooth <- dat2 %>% filter(Race2 == "Black" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, life.expectancy.birth.smoothed) %>%
                  rename(le.smoothed.black = life.expectancy.birth.smoothed)

WhiteSmooth <- dat2 %>% filter(Race2 == "White" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, life.expectancy.birth.smoothed) %>%
                  rename(le.smoothed.white = life.expectancy.birth.smoothed)

@@

BlackWhiteSmooth <- BlackWhiteSmooth %>% mutate(WBgap.s = le.smoothed.white - le.smoothed.black)

dim(BlackWhite)
BlackWhite3 <- merge(BlackWhite, BlackWhiteSmooth, by = c("State2", "Year3", "Sex2"))
dim(BlackWhite3)

BlackWhite <- BlackWhite3
rm(BlackWhite3, BlackSmooth, WhiteSmooth)
```

```{r calculate_age_decomp_smoothed}
#18 mins

list.age.decomp.tables.smoothed <- vector(mode = "list", length = dim(paired.ids)[1])

for(i in 1:dim(paired.ids)[1]) {
  white.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexWhite"])
  black.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexBlack"])

  if(length(white.index) != 0 & length(black.index) != 0) {
    list.age.decomp.tables.smoothed[[i]] <- le_age_decomp(life.table1 = as.data.frame(list.life.tables.smoothed[[black.index]]),
                                                          name.lt1 = "Black", 
                                                          life.table2 = as.data.frame(list.life.tables.smoothed[[white.index]]),
                                                          name.lt2 = "White", age.groups = "Age3")
  } else {
    list.age.decomp.tables.smoothed[[i]] <-"Couldn't compare life tables"
  }
  print(i)
}
```

```{r calculate_cod_decomp}
#takes 40 mins

list.cod.decomp.tables.smoothed <- vector(mode = "list", length = dim(paired.ids)[1])

system.time(
for(i in 1:dim(paired.ids)[1]) {
  white.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexWhite"])
  black.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexBlack"])

  if(length(white.index) != 0 & length(black.index) != 0) {
    ct.white <- subset(dat.clean, StateYearRaceSex == paired.ids[i, "StateYearRaceSexWhite"]) %>% 
                ungroup() %>%
                select(Age2, COD2, cod_prop_smoothed_deaths) %>%
                rename(prop.white = cod_prop_smoothed_deaths)

    ct.black <- subset(dat.clean, StateYearRaceSex == paired.ids[i, "StateYearRaceSexBlack"]) %>% 
                ungroup() %>%
                select(Age2, COD2, cod_prop_smoothed_deaths) %>%
                rename(prop.black = cod_prop_smoothed_deaths)
    ct.both <- merge(ct.white, ct.black, by = c("Age2", "COD2")) 
    ct.both <- ct.both %>% arrange(COD2, Age2)

    if(any(is.na(ct.both$prop.black)) == F & any(is.na(ct.both$prop.white)) == F) { #ensures that there are no ages with 0 deaths - for these COD prop == 0 and can't compute

        list.cod.decomp.tables.smoothed[[i]] <- cause_of_death_decomp(life.table1 = as.data.frame(list.life.tables.smoothed[[black.index]]),
                                                                      life.table2 = as.data.frame(list.life.tables.smoothed[[white.index]]), 
                                                                      decomp.table = as.data.frame(list.age.decomp.tables.smoothed[[i]]),
                                                                      cod.table = ct.both, 
                                                                      age.colname.cod.table = "Age2", 
                                                                      COD.colname.cod.table = "COD2", 
                                                                      prop1.colname.cod.table = "prop.black", 
                                                                      prop2.colname.cod.table = "prop.white")

    } else { list.cod.decomp.tables.smoothed[[i]] <- "Encountered age group with no deaths"

    }
  } else {

    list.cod.decomp.tables.smoothed[[i]] <- "Couldn't compare life tables"
  }

  print(i)
}
)
```

```{r  update_decomp_for_plotting}
#update the decomposition table for plotting purposes
for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables.smoothed[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables.smoothed[[i]][1] != "Encountered age group with no deaths") {

       list.cod.decomp.tables.smoothed[[i]] <- make_dataset_cod_plot(cod.decomp.table = list.cod.decomp.tables.smoothed[[i]],
                                                            age.groups = "Ages", 
                                                            cause.of.death = "Cause.of.death")  
    } 
  print(i)
}  
```

@@

```{r save_specified_files}
save(dat.clean, dat2, BlackWhite, BlackWhiteSmooth, paired.ids, list.age.decomp.tables, list.cod.decomp.tables,
     list.age.decomp.tables5, list.cod.decomp.tables5, list.age.decomp.tables9, list.cod.decomp.tables9,
     list.age.decomp.tables.smoothed, list.cod.decomp.tables.smoothed,
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/main_datasets.Rdata")
```
